We study how to build and---crucially---how to \emph{honestly evaluate} open-vocabulary
3D scene graphs from a handful of casually captured RGB views. Our pipeline keeps every
heavy model frozen and uses no per-scene optimization: feed-forward VGGT geometry
\cite{wang2025vggt} turns sparse images into world-point maps, an open-vocabulary
detector (OWLv2) \cite{minderer2023owlv2} proposes labeled object boxes, and these are
lifted to 3D and fused across views into a labeled object--relation graph. We make three
points. First, replacing the common class-agnostic-mask + per-patch CLIP
\cite{kirillov2023segment,radford2021clip} front-end with a genuine open-vocabulary
detector is what makes the labels real: on a desk scene the detector recovers the actual
objects (keyboard, monitors, book) where the mask+CLIP pipeline returned room-scale
category noise. Second, the dominant failure of the lifted system is \emph{over-counting}---one
physical object survives as several fused nodes---and a simple duplicate-aware fusion step
(merge same-label nodes whose 3D boxes overlap) fixes it, improving object-label F1 at every
view count with no losses on any scene and a clean five-of-five sweep at the denser views.
Third, a previously promising uncertainty-aware fusion
gate gives \emph{no} benefit once labels are real, across a full parameter sweep---a negative
result that a circular evaluation had masked. To ground these claims we contribute an
independent, human-verified sparse-view benchmark on TUM RGB-D \cite{sturm2012tumrgbd} whose
reference is authored from the frames rather than seeded from the model's own predictions, and
we document the circular-pseudo-reference pitfall that inflates scores. We hope the benchmark
and the de-circularization protocol prove useful beyond this system.
