\section{Discussion}
\label{sec:discussion}

Three findings of this paper pull in the same direction, and it is worth stating
plainly why. On an independent, human-verified reference, a one-line
duplicate-aware merge improves open-vocabulary object-label F1 at every view
count and on every scene, while a recalibrated, fully swept uncertainty-aware
fusion gate gives no improvement at all and in fact regresses. These two results
are not coincidental: they are two sides of the same mechanism. This section
explains that mechanism, draws the methodological lesson from the circular
evaluation that previously hid it, and is candid about the limits of the present
study.

\subsection{Why duplicate-aware merging helps and uncertainty does not}
\label{sec:disc-mechanism}

The dominant error of the lifted system is \emph{over-counting}: a single
physical object survives cross-view fusion as several nodes. The per-class
characterization makes this concrete. As the view count grows from three to ten,
the system's recall rises (micro-averaged $0.671 \rightarrow 0.808$) but its
precision falls in lockstep ($0.533 \rightarrow 0.322$), so net F1 is best at the
sparse end and degrades with more frames. The classes that drive the precision
collapse are exactly the ones that are \emph{always} found and then fragmented:
\texttt{monitor}, \texttt{desk}, \texttt{keyboard}, and \texttt{wall} all reach
recall $\approx 1.0$ while their precision falls toward $0.2$--$0.3$. More views
supply more correct detections of the same instance, and 3D fusion fails to
collapse them. The failure is over-segmentation of correct objects, not
mislabeling.

Read on this axis, the two interventions are opposites. Duplicate-aware fusion
merges same-label nodes whose axis-aligned 3D boxes overlap; it is a
\emph{merge}, and merging is the corrective move when the error is over-splitting.
The per-class effect is exactly as predicted: at eight views, precision jumps from
$0.38$ to $1.00$ on \texttt{monitor}, $0.19$ to $0.83$ on \texttt{desk}, and
$0.31$ to $0.62$ on \texttt{keyboard}, with recall preserved for nearly every
class. The net is a uniform F1 gain of $+0.068$, $+0.089$, $+0.145$, and $+0.146$
over the strongest non-merging fusion baseline at three, five, eight, and ten
views, and it is the only variant whose accuracy \emph{rises} as more views are
added --- because each added view supplies redundancy that the merge can absorb
rather than fragmentation that it cannot.

The uncertainty-aware gate acts on the same axis but in the wrong direction. It
\emph{tightens} the merge criterion for high-uncertainty pairs, i.e.\ it splits
more readily exactly where the system already over-splits. Against the
independent reference it never helps: across the entire weight sweep $w \in [0.1,
0.8]$ no setting beats the non-gated baseline, and the wired setting regresses by
$-0.014$, $-0.031$, $-0.056$, and $-0.057$ across the four view counts, losing on
all five scenes at five, eight, and ten views. This is not a tuning failure to be
patched away; it is the predictable consequence of suppressing merges in a regime
whose error is too few merges. We report it as a feature of the study, not an
embarrassment: knowing which lever points the wrong way is what lets us identify
the one that points the right way, and the contrast between the two variants is
the clearest evidence for the over-counting diagnosis itself.

\subsection{The circular-evaluation lesson}
\label{sec:disc-circularity}

This mechanism was invisible under the evaluation we --- like much of this
literature --- started with. When the reference graph is seeded from the model's
own dense-view prediction, the protocol is circular: a fusion method that simply
reproduces its ten-view output scores $1.0$ at the reference view \emph{by
construction}, and recall-raising behavior is rewarded regardless of whether it
corresponds to real objects. Under that prediction-seeded reference the
uncertainty gate appeared to be a sparse-view win ($+0.040$ and $+0.021$ at three
and five views) and the over-counting precision sink was masked by the
artificial dense-view ceiling. Re-grading the identical model outputs against an
\emph{independently authored} human reference --- written from the frames, never
from the predictions --- reverses the sign of the uncertainty result and removes
the dense-view artifact, exposing the precision collapse and the merging fix
underneath.

The methodological takeaway generalizes beyond this system. A reference seeded
from any single model's output does not measure recognition; it measures
agreement with that model, and it silently grants a perfect score to whatever
produced the seed. In sparse-view 3D scene understanding, where dense
ground truth is expensive and the temptation to bootstrap labels from a strong
model is real, this failure mode is easy to fall into and hard to notice, because
the inflated numbers look like success. We therefore treat the independent,
human-verified reference and the documentation of the circular-pseudo-reference
pitfall as a first-class contribution, and we recommend that any prediction-seeded
reference be reported alongside an independently authored one whenever the latter
is feasible. Concerns about evaluation construction and reproducibility in 3D
scene understanding have been raised before~\cite{gundersen2018reproducibility,kapoor2023leakage}; our point is the specific and easily
overlooked case where the reference is derived from the system under test.

\subsection{Limitations}
\label{sec:disc-limitations}

\paragraph{Evaluation scope.} The headline results are measured on $n = 5$
human-verified sequences. This is enough to establish the over-counting signature
and to show the merge wins on every scene at the dense views (with no losses at any
view count), but the sign-test floor at this sample
size is only $p = 0.062$, so per-scene win counts carry more weight than the
$p$-value. A roughly thirty-scene corroboration is in progress and will slot into
the same tables; we report the present numbers as the verified core and flag the
larger study as forthcoming. \todo{Insert $\sim$30-scene corroboration results
once the expansion run completes.}

\paragraph{Closely-packed same-class instances.} Box-IoU merging cannot separate
distinct instances of the same class whose axis-aligned 3D boxes overlap. A desk
holding two or three monitors side by side is the failure case: the merge collapses
them into one node, costing $-0.33$ recall on \texttt{monitor} and $-0.17$ on
\texttt{desk} at eight views. The net effect is still strongly positive because
the precision gain dwarfs the recall cost, but geometry alone is the wrong tool
for telling two adjacent same-class objects apart, and this is the one place the
merge does harm.

\paragraph{Static, ``thing''-centric, rule-based.} The pipeline assumes a static
incoherently. ``Stuff'' classes such as \texttt{wall}, \texttt{floor}, and
\texttt{ceiling} are treated as countable objects, which is a poor fit for amorphous
surfaces and inflates the over-counting they exhibit; they could reasonably be
reported separately or handled by a region rather than an instance model. Finally,
the spatial relations are inferred from centroid and distance rules. These are
inspectable and easy to audit, but they cannot express the support, containment,
functional, or activity relations studied in richer 3D scene graph
settings~\cite{armeni2019scenegraph,wald2020threedssg,gu2024conceptgraphs}.

\subsection{Future work}
\label{sec:disc-future}

The cleanest next step follows directly from the merge's one failure mode.
Box-IoU over-merges closely-packed same-class instances; appearance- or
feature-based instance separation --- splitting a merged group back apart when its
members are visually distinct --- is the natural remedy. This is also, notably,
where the recall-oriented signal that the uncertainty gate tried and failed to add
could finally pay off: not as a \emph{merge gate} that prevents fusion in the
over-counting regime, but as a \emph{split decision} applied to an
already-merged group, exactly the direction in which a recall signal helps rather
than hurts. The negative uncertainty result thus does not retire the idea; it
relocates it to the one operation where it is mechanistically appropriate. Beyond
this, replacing the rule-based predicates with a learned relation module --- once
the independently authored annotations provide sufficient object-relation
supervision --- and reporting ``stuff'' classes under a region model are the
obvious extensions.

\section{Conclusion}
\label{sec:conclusion}

We presented a detector-grounded open-vocabulary 3D scene-graph system that turns
a handful of casual RGB views into an inspectable, language-grounded
object-relation graph with all heavy models frozen and no per-scene optimization:
an open-vocabulary detector (OWLv2)~\cite{minderer2023owlv2} proposes labeled
boxes, feed-forward VGGT~\cite{wang2025vggt} geometry lifts them to 3D, and a
cross-view fusion step assembles the graph. On an independent, human-verified
TUM RGB-D~\cite{sturm2012tumrgbd} benchmark we showed three things. The dominant
error is over-counting, and a one-line duplicate-aware merge corrects it,
improving object-label F1 by $+0.07$ to $+0.15$ at every view count, with no losses
on any scene and a clean sweep at the dense views. A recalibrated, fully swept uncertainty-aware fusion gate --- the move in
the opposite direction --- gives no benefit, a negative result that a circular,
prediction-seeded reference had previously disguised as a win. And the
de-circularized reference that exposed both, together with the documented pitfall
that produced the original illusion, is the methodological core we most hope
outlasts this particular system. What actually helps in sparse-view 3D scene
graphs, on honest labels, turned out to be the simplest move pointed in the right
direction.
