\section{Experiments}
\label{sec:experiments}

We evaluate two questions that the reframed thesis turns on: (i) what actually
helps when building open-vocabulary 3D scene graphs from a handful of RGB views,
and (ii) how to measure that honestly. The second question is not incidental.
The system we study was originally reported as a sparse-view win for
uncertainty-aware fusion; that conclusion was an artifact of a circular
evaluation, and exposing it required first building a reference that the system
could not grade itself against. We therefore describe the benchmark and its
de-circularization in detail before reporting the variant comparison
(Section~\ref{sec:results}).

\subsection{Benchmark Scenes and Sparse-View Protocol}
\label{sec:exp_scenes}

We benchmark on public TUM RGB-D indoor sequences~\cite{sturm2012tumrgbd}. The
benchmark spans \textbf{28 object-rich sequences} across \texttt{freiburg1}--\texttt{freiburg3}:
the five paper-subset scenes in Table~\ref{tab:tum_rgbd_paper_subset_plan} (which carry the
human-verified gold reference) plus 23 additional sequences; the full list is in
\texttt{configs/datasets/tum\_rgbd\_expanded.json}. Each sequence contributes ten
selected RGB frames (stride~10). Although the source benchmark is RGB-D, our
pipeline consumes only RGB; we use this dataset because it provides public,
recognizable indoor scenes with reproducible sequence identifiers.

\input{tables/tum_rgbd_paper_subset_plan}

For each scene we evaluate four input-view counts, $N \in \{3, 5, 8, 10\}$,
giving a 28-scene by four-view-count matrix (112 graph outputs per variant). The benchmark
runner creates one output directory per (scene, $N$) and executes the full
pipeline along an identical code path: VGGT geometry, OWLv2 detection,
descriptor extraction, 3D lifting, cross-view fusion, open-vocabulary labeling,
and metric export. The driver is \texttt{scripts/run\_owlv2\_benchmark.sh}.
Reporting accuracy as a function of $N$ lets us read off the sparse-view
behavior of each variant directly, rather than a single dense-view number.

\subsection{An Independently-Authored Reference}
\label{sec:exp_reference}

The central evaluation problem in this setting is that there is no
ground-truth open-vocabulary object inventory for TUM RGB-D scenes, so a
reference must be constructed. The tempting shortcut --- and the one the
original version of this work took --- is to seed the reference from the
system's own dense-view prediction (the 10-view graph) and then lightly check
it. We argue this is unsound and replace it with a reference authored
\emph{from the frames}.

\paragraph{Why the prediction-seeded reference is circular.}
If the reference object multiset is initialized from a variant's own 10-view
output, that variant scores precision~$=$~recall~$=$~$1.0$ at the 10-view
``reference'' point \emph{by construction}, and every sparser run is graded for
how faithfully it reproduces the same variant's splitting and labeling pattern
rather than for agreement with the scene. A method that keeps more uncertain
proposals unmerged then looks better simply because it reproduces more of the
fragments the reference inherited from the seeding variant. Under exactly this
prediction-seeded reference, the uncertainty-aware variant appeared to be a
sparse-view win ($+0.040$/$+0.021$ at $N{=}3$/$5$) and the fusion baseline hit
$1.0$ at $N{=}10$ by construction. Both effects are evaluation artifacts, not
properties of the scenes.

\paragraph{Construction of the independent reference.}
We instead enumerate, for each scene, an object multiset built only
from the ten raw RGB frames and \emph{independent} of the
detection/descriptor/geometry/fusion stack under test. Each scene is drafted in
one pass and then re-examined in a second, adversarial pass (an explicit
attempt to find missed or spurious objects) over the same frames. For the five
paper-subset scenes the two-pass draft is additionally \textbf{human-verified}
against the frames, removing ambiguous detections (e.g.\ a window glimpsed in
\texttt{room}/\texttt{desk}/\texttt{desk2}; cup/bottle/picture in \texttt{desk2});
these five form the gold core (\texttt{configs/evaluation/independent\_labels.json}).
The 23 additional scenes use the same two-pass draft \emph{without} human
verification (\texttt{independent\_labels\_expanded.json}) and serve as scale
corroboration; we show both main findings also hold on the human-verified five
alone. Per-scene provenance is logged in
\texttt{docs/independent\_reference\_worklist.md}. Because this reference fixes
the true object inventory without reference to any variant's output, no variant
can score $1.0$ at $N{=}10$ by construction, and over-splitting is correctly
penalized rather than rewarded. The reference is a scene-level multiset over the
ten frames; sparse-view predictions are recall-limited against it by
construction, which depresses absolute F1 but applies uniformly across variants
and so leaves the comparison fair.

\subsection{Variants Compared}
\label{sec:exp_variants}

We hold the front-end (OWLv2 detections lifted by VGGT geometry) and the
reference fixed, and compare seven variants that differ only in how per-view
detections become 3D nodes and how those nodes are fused across views:

\begin{itemize}
  \item \textbf{2d-only} --- per-view OWLv2 labels pooled without 3D lifting or
        cross-view fusion (an image-level upper reference on labels alone).
  \item \textbf{geometry-only} --- detections lifted to 3D via VGGT and pooled,
        with no semantic cross-view association.
  \item \textbf{semantic-lifting} --- detections lifted and associated by
        appearance descriptors without the geometric fusion gate.
  \item \textbf{fixed-shrink (uninformed control)} --- the fusion gate is
        modulated by a fixed, scene-agnostic shrink factor; an
        uncertainty-blind control that isolates whether the
        \emph{uncertainty signal}, rather than merely the act of tightening the
        gate, carries any benefit.
  \item \textbf{graph-fusion (baseline)} --- the standard cross-view fusion of
        Section~\ref{sec:method} with no uncertainty modulation and no
        duplicate handling.
  \item \textbf{proposed (uncertainty)} --- graph-fusion with the
        rank-normalized uncertainty-aware merge gate (default weight
        $w{=}0.3$); evaluated as an ablation rather than the headline method,
        and additionally swept over $w \in \{0.1, 0.2, 0.3, 0.5, 0.8\}$.
  \item \textbf{graph-fusion-dedup (ours)} --- graph-fusion followed by a single
        post-fusion duplicate-instance merge: same-label fused nodes whose
        axis-aligned 3D boxes overlap (IoU~$>\tau$, default $\tau{=}0.1$) are
        unioned and re-fused (\texttt{merge\_duplicate\_instances} in
        \texttt{graph\_builder.py}). This targets the measured over-counting
        failure (Section~\ref{sec:results}) and is the opposite move from the
        uncertainty gate, which \emph{tightens} merging.
\end{itemize}

The pairing is deliberate: \emph{fixed-shrink} versus \emph{proposed} asks
whether the uncertainty signal is informative beyond an uninformed gate
adjustment, and \emph{graph-fusion} versus \emph{graph-fusion-dedup} asks
whether merging \emph{more} (rather than less) is the productive direction.

\subsection{Metric}
\label{sec:exp_metric}

Our primary metric is \texttt{object\_label\_f1}: the multiset precision,
recall, and F1 of fused-node labels against the independent reference multiset,
averaged over the 28 scenes (5 human-verified, 23 two-pass VLM-drafted). Treating
labels as a multiset (rather than a set) makes the metric sensitive to
over-counting --- a scene with one physical monitor reported as four fused monitor
nodes is penalized on precision --- which is exactly the failure mode the reframed
contribution targets. Per-scene wins/losses are reported alongside the means, and
significance is summarized with a two-sided sign-test, which over the 28 scenes
reaches $p<0.001$ for both main findings. Evaluation is run by
\texttt{scripts/evaluate\_sparse\_view\_annotations.py} and aggregated by
\texttt{scripts/aggregate\_variant\_f1.py}; raw per-(scene, $N$, variant) scores
are in \texttt{results/benchmark\_owlv2\_expanded/variant\_independent\_metrics.csv}.

\subsection{Implementation Details}
\label{sec:exp_impl}

All heavy models are frozen and run off-the-shelf; there is no per-scene
optimization or fine-tuning. The open-vocabulary detector is OWLv2
(\texttt{google/owlv2-base-patch16-ensemble})~\cite{minderer2023owlv2}, the
successor to OWL-ViT~\cite{minderer2022owlvit}, with a detection score threshold
of $0.2$ and per-class non-maximum suppression at IoU~$0.5$; detection boxes
prompt SAM~\cite{kirillov2023segment} for masks. 3D geometry is the feed-forward
VGGT transformer~\cite{wang2025vggt}, consuming RGB frames only. Appearance
descriptors for cross-view association use CLIP~\cite{radford2021clip} and
DINOv2~\cite{oquab2023dinov2}. VGGT geometry runs on GPU; detection, lifting,
fusion, and evaluation run on CPU. We fix all thresholds across scenes and view
counts (no per-scene tuning), and the duplicate-aware merge uses a single global
IoU threshold $\tau{=}0.1$ that we show is robust across $\tau \in [0, 0.2]$
(Section~\ref{sec:results}).

\subsection{Gold Core, Drafted Corroboration, and Dynamic Stress}
\label{sec:exp_expansion}

The 28-scene benchmark pairs a small \emph{human-verified} core (the five paper-subset scenes)
with 23 two-pass VLM-drafted scenes for statistical power. Both main findings reach $p<0.001$
over the 28 scenes and reproduce on the five human-verified scenes alone
(Section~\ref{sec:results}), so they do not depend on the drafted labels; fully human-verifying
the 23 is the natural next step. Six sequences are \emph{dynamic} (a moving person:
\texttt{desk\_with\_person}, \texttt{sitting\_*}, \texttt{walking\_*}), which violates VGGT's
static multi-view assumption; we report them as a stress subset. The duplicate-aware gain holds
in both groups --- $+0.082$ to $+0.171$ on the 22 static scenes and $+0.060$ to $+0.116$ on the
6 dynamic scenes --- so the contribution is robust to scene dynamics as well as to scale.
