\section{Introduction}

\input{figures/fig_qual}

Reconstructing a scene and explaining a scene are related but different goals. Modern
geometry pipelines can recover camera motion, depth, points, surfaces, or radiance
fields, yet downstream applications often require object-level statements: which
pixels across views depict the same object, where that object is in 3D, and how it is
spatially related to other objects. A robot or mixed-reality system, for example,
benefits more from a statement such as ``the monitor is on the desk'' than from an
unlabeled collection of points. This need is especially acute when the input is small:
a few casual RGB views rather than a calibrated RGB-D stream, dense video, or an
already optimized map.

Classical structure-from-motion and SLAM systems remain powerful when overlap,
calibration, and optimization time are available~\cite{schonberger2016sfm,campos2021orbslam3}.
Recent feed-forward geometry models have changed the starting point by predicting
camera parameters, depth, and point maps directly from one or more
images~\cite{wang2024dust3r,leroy2024mast3r,wang2025vggt}. At the same time, 2D
foundation models provide strong open-vocabulary detections, language-aligned
embeddings, class-agnostic masks, and robust self-supervised descriptors through
detectors such as OWL-ViT and OWLv2, and through CLIP, SAM, and
DINOv2~\cite{minderer2022owlvit,minderer2023owlv2,radford2021clip,kirillov2023segment,oquab2023dinov2}.
These components make sparse-view semantic 3D reasoning plausible, but they do not by
themselves define an object-level graph. Per-image detections are local, repeated
instances may have similar descriptors, and feed-forward geometry is least reliable
near occlusions, thin structures, textureless regions, and large viewpoint changes.

We approach sparse-view scene graph construction as evidence-preserving fusion over
lifted 2D object proposals, keeping all heavy perception models frozen and using no
per-scene optimization. VGGT world-point maps serve as a 2D-to-3D lookup table; an
open-vocabulary detector proposes labeled object boxes; these are lifted into candidate
3D objects, merged across views into a labeled object-relation graph, and annotated
with spatial predicates. This design keeps the expensive perception stage fixed while
making the graph reasoning stage explicit, auditable, and easy to ablate. It also
differs from open-vocabulary 3D graph systems that assume posed RGB-D video, a dense
point cloud, or a prebuilt map~\cite{gu2024conceptgraphs,jatavallabhula2023conceptfusion}:
our target is a lower-input regime of sparse RGB views processed by feed-forward
geometry.

A central lesson of this work is methodological. An earlier version of this system
graded its predicted scene graphs against a pseudo-reference \emph{seeded from the
model's own dense-view output}. That protocol is circular: a fusion method that simply
reproduces its dense-view graph scores $1.0$ at the reference view by construction, and
the apparent gains it reports need not survive an honest test. When we re-grade against
an \emph{independently authored}, human-verified reference --- written from the frames
rather than from any model prediction --- two things change. First, a front-end that
labeled image patches with CLIP turns out to recover almost none of the real objects:
on a desk scene it returns room-scale category noise (``floor'', ``curtain'', ``bed'')
rather than the monitors, keyboard, and books that are actually present, and its
headline label accuracy was an artifact of grading CLIP against CLIP. Replacing the
patch-labeling front-end with a genuine open-vocabulary \emph{detector} (OWLv2) makes
the labels real: the detector recovers the actual objects on every frame. Second, and
more cautionary, a previously promising uncertainty-aware fusion gate --- which our
earlier circular protocol reported as a sparse-view win --- shows \emph{no} benefit
once labels are real, across a full weight sweep. We report this negative result as a
feature, not a footnote: it is exactly the kind of conclusion a de-circularized
evaluation is meant to surface.

With honest labels in hand, the dominant remaining error is over-counting: one physical
object survives cross-view fusion as several same-label nodes (Figure~\ref{fig:qual}). The fix
is deliberately simple --- a duplicate-aware merge of same-label nodes whose axis-aligned 3D boxes
overlap. This step improves object-label F1 at every view count and is the only variant
whose accuracy \emph{rises} with more views, precisely because it converts the recall
that extra views provide into precision rather than into duplicates. Notably, this is
the opposite move from the uncertainty gate (which tends to over-split), and it
explains why merging helps where the gate does not.

We evaluate on 28 public TUM RGB-D sequences under 3/5/8/10-view
protocols~\cite{sturm2012tumrgbd}, against an independently authored reference (5 scenes
human-verified, 23 drafted by two independent passes). The duplicate-aware variant improves
object-label F1 over the no-uncertainty graph-fusion baseline by $+0.06$ to $+0.12$ across view
counts and wins on $22$--$27$ of the $28$ scenes ($p<0.001$), while the recalibrated uncertainty
gate is consistently \emph{worse} than the baseline and an uninformed control. The gain holds on
the $5$ human-verified scenes alone and on both static and dynamic (moving-person) sequences.

Our contributions are:
\begin{itemize}
    \item \textbf{A detector-grounded open-vocabulary sparse-view 3D scene-graph system.}
    An open-vocabulary detector (OWLv2) proposes labeled boxes, VGGT geometry lifts them
    to 3D, and a cross-view fusion stage produces a labeled object-relation graph with
    spatial predicates --- all models frozen, no per-scene optimization. Detector-grounded
    labels recover real objects where a mask-plus-CLIP-patch front-end does not.
    \item \textbf{Duplicate-aware cross-view fusion.} A post-fusion merge of same-label
    nodes whose axis-aligned 3D boxes overlap ($\mathrm{IoU} > \tau$), targeting the
    measured over-counting failure. It improves object-label F1 at every view count and
    is the only variant whose accuracy rises with more views.
    \item \textbf{An independent, human-verified sparse-view benchmark and a
    de-circularization protocol on TUM RGB-D.} We document how a prediction-seeded
    pseudo-reference inflates scores (graph fusion scores $1.0$ at the reference view by
    construction) and show that an independently authored reference removes the artifact.
    \item \textbf{A rigorous negative result on uncertainty-aware fusion.} Recalibrated
    and swept over weights, the uncertainty gate does not improve object accuracy on real
    labels; we analyze why (it lifts recall but over-splits, costing precision), which
    also explains why the opposite move --- duplicate-aware merging --- works.
\end{itemize}
