\section{Method}
\label{sec:method}

\subsection{Problem Setup}

Given a sparse set of RGB views $\mathcal{I}=\{I_1,\ldots,I_V\}$, the goal is to
produce an open-vocabulary 3D scene graph $\mathcal{G}=(\mathcal{O},\mathcal{R})$.
Each object node $o_i \in \mathcal{O}$ stores a 3D point set, centroid, an
axis-aligned 3D bounding box, an open-vocabulary label, view support, a feature
descriptor, and its source detections. Each relation edge $r_{ij} \in \mathcal{R}$
stores a spatial predicate between two object nodes. The method assumes neither depth
images, externally supplied camera poses, nor a prebuilt mesh: it derives geometry and
labeled object proposals from frozen feed-forward models, lifts the proposals to a
common 3D frame, and fuses them into objects with no per-scene optimization.
Figure~\ref{fig:method_pipeline} summarizes the data flow.

\input{figures/fig_pipeline}

\subsection{Overview}

The pipeline has five stages, all built on frozen models. (i)~VGGT predicts camera
parameters, depth maps, world-point maps, and confidence maps for the sparse
views~\cite{wang2025vggt}. (ii)~An open-vocabulary \emph{detector} (OWLv2) proposes
labeled object boxes from a text query set, so each proposal already carries a real
object label~\cite{minderer2023owlv2}. (iii)~Each detection is lifted into the VGGT
world-point map to form a 3D candidate fragment. (iv)~Candidates from different views
are fused into object nodes, after which a duplicate-aware pass merges over-counted
nodes that correspond to the same physical object. (v)~Fused nodes are connected by
rule-based spatial relations. The exported graph keeps the predicted labels and
relations alongside the evidence that produced them, so it can be inspected or turned
into annotation-review packets.

A central design change relative to a common open-vocabulary recipe deserves emphasis
up front. A widely used front-end pairs class-agnostic masks~\cite{kirillov2023segment}
with per-region CLIP scoring~\cite{radford2021clip} to label regions
post~hoc~\cite{jatavallabhula2023conceptfusion,gu2024conceptgraphs,takmaz2023openmask3d}.
On blurry, low-resolution casual frames this front-end produced surface and background
fragments that CLIP then labeled with room-scale categories, yielding labels
disconnected from the actual objects (Section~\ref{sec:experiments}). We therefore
replace it with a genuine open-vocabulary detector whose box \emph{is} the labeled
object hypothesis. Everything downstream of the front-end --- lifting, fusion, and
relations --- is kept unchanged, which lets us attribute the resulting accuracy change
to the front-end rather than to a re-tuned geometry stack.

\subsection{Sparse-View Geometry}
\label{sec:geometry}

We run VGGT on the selected sparse views and store camera intrinsics, extrinsics, depth
maps, world-point maps, and confidence maps~\cite{wang2025vggt}. The world-point map for
each view acts as a 2D-to-3D lookup table: object proposals are lifted directly from
image coordinates into a common 3D frame without first reconstructing a mesh, running
bundle adjustment, or optimizing a radiance-field or Gaussian-Splatting
representation~\cite{schonberger2016sfm,mildenhall2020nerf,kerbl2023gaussian}. For view
$v$ we denote the world-point map by $P_v(u) \in \mathbb{R}^3$ and the confidence map by
$C_v(u)$ at image coordinate $u$.

\subsection{Open-Vocabulary Detection Front-End}
\label{sec:owlv2}

\input{figures/fig_detections}

For each view we run OWLv2~\cite{minderer2023owlv2}, an open-vocabulary detector that
takes a set of free-form text queries and returns boxes with per-box class scores. The
queries are the indoor object vocabulary of the benchmark, so the detector --- not a
post~hoc patch-classification step --- carries the open-vocabulary label. For detection
$k$ in view $v$ we store its box, score $s_{v,k}$, and predicted label
$\ell_{v,k}$. We use the \texttt{owlv2-base-patch16-ensemble} model with a detection
threshold of $0.2$ and per-class non-maximum suppression at box-IoU $0.5$; the NMS
mainly cleans the large surface classes (e.g.\ a desk) that otherwise accrue many
overlapping boxes. We deliberately do \emph{not} over-tighten the 2D threshold: a
physical object may receive several overlapping boxes per frame, but overlapping boxes
on the same object lift to nearly the same 3D points and are intended to collapse
downstream. Optionally, each detection box prompts a frozen
SAM~\cite{kirillov2023segment} to produce a tight instance mask $M_{v,k}$ so that
lifting selects only object pixels; without a mask, lifting falls back to the box
rectangle. When language-aligned and self-supervised appearance descriptors are
available, we additionally extract CLIP and DINOv2 features for each detection and store
the normalized concatenation~\cite{radford2021clip,oquab2023dinov2}; these are used only
for fusion gating and provenance, never for labeling.

\subsection{3D Lifting}
\label{sec:lifting}

Each detection's region (its instance mask $M_{v,k}$, or its box if no mask is
available) is resized to the VGGT geometry resolution and used to select world points
from the corresponding view; non-finite points and points below the confidence
threshold $\tau_c$ are discarded:
\begin{equation}
    X_{v,k} = \{P_v(u) \mid u \in M_{v,k},\; C_v(u) \geq \tau_c\}.
\end{equation}
Detections with no valid lifted points are rejected. Each surviving candidate node
stores its sampled 3D point set $X_{v,k}$, centroid $\mu_{v,k}$, an axis-aligned 3D
bounding box, the detector label $\ell_{v,k}$ and score $s_{v,k}$, and (when available)
the feature $f_{v,k}$.

\subsection{Cross-View Fusion}
\label{sec:fusion}

Candidate nodes are fused into objects by building an association graph over the lifted
detections. By default we disallow same-view association edges, so a connected component
represents cross-view support for one object rather than a grouping of detections within
a single image. An edge connects two candidates when their centroids are close in 3D
and, when valid descriptors exist, their cosine similarity exceeds a feature threshold:
\begin{equation}
    \|\mu_a-\mu_b\|_2 \leq \epsilon_g
    \quad\text{and}\quad
    \cos(f_a,f_b) \geq \epsilon_f .
\end{equation}
The feature condition is omitted for candidates without valid descriptors. Connected
components define fused object identities; for each component, points are pooled and
sampled, descriptors are averaged and renormalized, and the node label is set by a
score-weighted majority vote over the constituent detector labels $\{\ell_{v,k}\}$
(ties broken by count). This separates view-local evidence from object identity while
preserving the detection provenance needed for inspection and annotation.

\subsection{Duplicate-Aware Fusion}
\label{sec:dedup}

Cross-view fusion alone leaves the dominant precision failure unaddressed.
Characterizing the fused system against an independent reference
(Section~\ref{sec:experiments}) shows that the precision loss is almost entirely
\emph{over-counting of correctly found objects}, not mislabeling: as more views are
added the detector finds more of the true objects (recall rises), but each physical
object also accumulates more detections that fusion fails to collapse into a single
node (precision falls). Two structural reasons make this expected. First, the detector
over-fires, emitting several overlapping boxes per object per frame. Second, by design
the association graph forbids same-view edges, so duplicate detections of one object
within a frame can never be joined by cross-view fusion. The result is that a single
desk or monitor survives as several fused nodes. Over-counting --- not the labels
themselves --- is therefore the lever to pull, and the corrective move is \emph{merging},
not splitting.

We add a single post-fusion pass, \textsc{Dedup}, that merges fused nodes of the same
label whose axis-aligned 3D boxes overlap. Let $\mathcal{O}$ be the fused nodes from
Section~\ref{sec:fusion}. For a node $o$ with point set $X_o$, let
$B_o=[\,b^{\min}_o,\,b^{\max}_o\,]$ be its axis-aligned 3D bounding box, with
$b^{\min}_o=\min X_o$ and $b^{\max}_o=\max X_o$ taken coordinate-wise. For two boxes the
axis-aligned 3D intersection-over-union is
\begin{equation}
    \mathrm{IoU}_{3D}(B_a,B_b)=
    \frac{\mathrm{vol}\!\left(B_a\cap B_b\right)}
         {\mathrm{vol}(B_a)+\mathrm{vol}(B_b)-\mathrm{vol}\!\left(B_a\cap B_b\right)},
\end{equation}
where $\mathrm{vol}(B_a\cap B_b)=\prod_{d=1}^{3}\max\!\bigl(0,\,
\min(b^{\max}_{a,d},b^{\max}_{b,d})-\max(b^{\min}_{a,d},b^{\min}_{b,d})\bigr)$. We
partition $\mathcal{O}$ by label. Within each label group we form a merge graph whose
vertices are the nodes and whose edges connect pairs with
$\mathrm{IoU}_{3D}(B_a,B_b)>\tau$, and we take its connected components; each component
is collapsed into one object by re-running the same pooling, descriptor averaging, and
label vote used in fusion. Nodes with no label, and singleton groups, are passed through
unchanged. Concretely, with $G_\ell=\{o\in\mathcal{O}\,:\,\mathrm{label}(o)=\ell\}$:
\begin{equation}
    \textsc{Dedup}(\mathcal{O};\tau)=
    \bigcup_{\ell}\;\Bigl\{\,\textsc{Fuse}(c)\;:\;
    c\in\mathrm{CC}\bigl(G_\ell,\;\{(a,b):\mathrm{IoU}_{3D}(B_a,B_b)>\tau\}\bigr)\Bigr\},
\end{equation}
where $\mathrm{CC}(\cdot)$ denotes connected components and $\textsc{Fuse}(\cdot)$
re-fuses a component into a single node (the identity on singletons). Restricting merges
to same-label nodes keeps genuinely distinct objects of different classes separate, and
requiring box overlap keeps spatially separated instances of the \emph{same} class
separate. The default threshold is $\tau=0.1$; the step is robust across the low-IoU
range and the IoU criterion alone outperforms adding a centroid-distance term
(Section~\ref{sec:experiments}). We use 3D-box IoU rather than mask or point-set overlap
because the lifted boxes are exactly the geometry already attached to each node, making
the test parameter-free beyond $\tau$. The one structural cost is that closely packed
same-class instances (e.g.\ two monitors on one desk) can have overlapping boxes and be
over-merged; we quantify this trade-off and discuss appearance-based instance splitting
as the natural remedy in Sections~\ref{sec:experiments} and~\ref{sec:discussion}.

\subsection{Spatial Relations}
\label{sec:relations}

Spatial relations are inferred from fused centroids using distance and relative-position
rules. The relation vocabulary includes \emph{near}, \emph{left\_of}, \emph{right\_of},
\emph{above}, \emph{below}, \emph{in\_front\_of}, and \emph{behind}. A \emph{near}
relation is emitted for object pairs within a distance threshold; otherwise the vertical
displacement or the dominant horizontal coordinate determines the directional predicate,
with an optional maximum relation distance suppressing very distant pairs. The exported
graph preserves both the symbolic fields and the numeric support, so the same object
identities, point sets, labels, and provenance can be reused by manual review or by a
future learned relation head.

\subsection{An Ablated Variant: Uncertainty-Aware Fusion}
\label{sec:uncertainty}

A natural hypothesis --- and the headline of an earlier version of this system --- is
that fusion should be \emph{uncertainty-aware}: each lifted candidate can be assigned an
uncertainty
\begin{equation}
    U_{v,k} = \lambda\,(1-\bar{C}_{v,k}) + (1-\lambda)\,
    \mathrm{clip}\!\left(\tfrac{\sigma(X_{v,k})}{s_0},0,1\right),
\end{equation}
combining VGGT point confidence $\bar{C}_{v,k}$ with the spatial spread
$\sigma(X_{v,k})$ of the lifted points (in our runs $\lambda=0.7$, $s_0=1$), and the
fusion gate is then \emph{tightened} for uncertain pairs --- effectively splitting
ambiguous candidates apart. We retain this as an ablated variant rather than as part of
the core method, because (i)~the failure analysis above shows the system over-counts, so
the productive move is to merge more, not to split more, and a gate that splits uncertain
pairs acts on the wrong axis; and (ii)~once labels are real and the evaluation is
de-circularized, the variant gives no benefit over plain cross-view fusion across a full
weight sweep (Section~\ref{sec:experiments}). We report it as a deliberate, rigorous
negative result. The uncertainty score itself is still exported with each node, since it
is potentially useful for the opposite operation --- guiding an appearance-based
\emph{split} of over-merged groups --- which we leave to future work
(Section~\ref{sec:discussion}).
