\section{Related Work}
\label{sec:related_work}

\paragraph{Feed-forward sparse-view geometry.}
Multi-view reconstruction and SLAM classically rely on correspondence, triangulation,
bundle adjustment, and calibrated or densely overlapping observations. COLMAP remains a
standard reference for incremental structure from motion~\cite{schonberger2016sfm},
while ORB-SLAM3 demonstrates the strength of optimized visual and visual-inertial SLAM
for coherent streams~\cite{campos2021orbslam3}. Feed-forward 3D models reduce this
dependence on per-scene optimization. DUSt3R casts reconstruction as point-map
prediction from unconstrained images~\cite{wang2024dust3r}; MASt3R adds 3D-aware dense
matching~\cite{leroy2024mast3r}; and VGGT predicts camera parameters, depth, point
maps, confidence, and tracks from one or many views~\cite{wang2025vggt}. These models
provide the geometric substrate for our work. We use VGGT as a frozen source of
pixel-aligned 3D evidence, but our output is not a reconstruction: it is an
object-relation graph grounded in that geometry.

\paragraph{Open-vocabulary detection as a labeling front-end.}
The semantics of a lifted scene graph are only as good as the 2D labels it ingests.
A common recipe pairs class-agnostic masks from SAM~\cite{kirillov2023segment} with
per-region CLIP embeddings~\cite{radford2021clip} matched against a text vocabulary,
optionally backed by DINOv2 appearance descriptors~\cite{oquab2023dinov2}. This
mask-then-embed route is attractive because every component is frozen, but it pushes the
naming burden onto whole-image-trained CLIP applied to cropped regions, which tends to
return scene- or room-scale category guesses rather than the individual objects present.
A separate line of work performs \emph{open-vocabulary detection} directly: the model
localizes \emph{and} names objects against an open text vocabulary in one pass. OWL-ViT
attaches lightweight detection heads to a contrastively pre-trained image--text
backbone and queries it with class-name embeddings~\cite{minderer2022owlvit}; OWLv2
scales this recipe with self-training on web image--text data and improves rare-category
localization~\cite{minderer2023owlv2}; Grounding DINO marries a DINO detector with
grounded language pre-training for open-set detection from text
prompts~\cite{liu2024groundingdino}; and Detic broadens detector vocabularies by
training on image-classification data~\cite{zhou2022detic}. We adopt an
open-vocabulary detector (OWLv2) as the front-end of our system precisely because it
emits localized, named instances, which is the property the lifting and fusion stages
need. To our knowledge this family has been used mainly for 2D benchmarks; we study what
it buys a sparse-view 3D scene-graph pipeline relative to the mask-then-embed
alternative.

\paragraph{Open-vocabulary 3D semantics.}
In 3D, open-vocabulary scene understanding has been explored by lifting, distilling,
or optimizing foundation-model features into point clouds, maps, masks, and neural
scene representations. NeRF introduced continuous radiance fields for view
synthesis~\cite{mildenhall2020nerf}; LERF embeds language features into radiance
fields for open-ended 3D queries~\cite{kerr2023lerf}; and 3D Gaussian Splatting has
become a common explicit substrate for fast scene representations~\cite{kerbl2023gaussian}.
OpenScene learns open-vocabulary point features aligned with image and text
embeddings~\cite{peng2023openscene}; ConceptFusion builds queryable 3D maps by fusing
pixel-aligned open-set features~\cite{jatavallabhula2023conceptfusion}; OpenMask3D
aggregates CLIP features over class-agnostic 3D instance masks~\cite{takmaz2023openmask3d};
CUA-O3D studies cross-modal and uncertainty-aware agglomeration for open-vocabulary
3D scene understanding~\cite{li2025cuao3d}; and OpenSplat3D performs open-vocabulary
3D instance segmentation in a Gaussian Splatting representation~\cite{piekenbrinck2025opensplat3d}.
These methods motivate foundation-feature fusion in 3D. Our contribution is
complementary: we construct explicit object-relation graphs from sparse RGB views
without assuming a completed 3D map or optimizing a per-scene radiance field. Notably,
uncertainty-aware aggregation has been proposed as a way to fuse multi-view evidence in
this setting~\cite{li2025cuao3d}; we revisit that idea on real detector labels under an
independent reference and report a negative result (Section~\ref{sec:results}).

\paragraph{3D scene graphs.}
3D scene graphs organize object instances and their relationships as structured
representations for reasoning about scenes. Early work proposed unified graphs spanning
objects, spaces, cameras, and semantic or spatial relationships~\cite{armeni2019scenegraph}.
3DSSG introduced a large-scale dataset and learned prediction framework for semantic
scene graphs from 3D indoor reconstructions~\cite{wald2020threedssg}. RGB-D resources
such as ScanNet further show the value of richly annotated reconstructed scenes for
indoor understanding~\cite{dai2017scannet}. ConceptGraphs is closest to our output
space: it builds open-vocabulary 3D scene graphs from posed RGB-D sequences by fusing
2D foundation-model outputs into object nodes and relations~\cite{gu2024conceptgraphs}.
We share the object-relation motivation, but relax the input assumption. Instead of
requiring a depth stream, posed RGB-D map, or completed 3D reconstruction, we lift
sparse-view 2D proposals through feed-forward VGGT point maps and build object
identities through cross-view geometric and semantic fusion. A recurring failure mode
of any such lifting-and-fusion pipeline is \emph{over-counting} -- one physical object
surviving as several fused nodes -- which motivates the duplicate-aware merge we study.

\paragraph{Honest evaluation and reproducibility in 3D scene understanding.}
Evaluating open-vocabulary 3D scene graphs is itself hard: there is rarely a clean,
instance-level ground truth for a handful of casual RGB views, so it is tempting to
score a method against a reference derived from the system's own dense-view output.
This couples the reference to the predictor and produces a \emph{circular} evaluation in
which a method can attain a perfect score at the reference view by construction, masking
real differences between variants. Such reference--predictor coupling is a form of
information leakage, which has been identified as a widespread driver of the
reproducibility crisis in machine-learning-based
science~\cite{kapoor2023leakage,gundersen2018reproducibility}. The broader 3D community
has responded by curating carefully controlled benchmarks with independent
ground truth -- e.g.\ the Replica dataset of high-fidelity reconstructed indoor
scenes~\cite{straub2019replica} and the long-standing TUM RGB-D benchmark with
externally measured trajectories~\cite{sturm2012tumrgbd}. In that spirit, we contribute
an independent, human-verified sparse-view reference authored from the frames rather
than seeded from the model's predictions, document the circular-pseudo-reference pitfall
explicitly, and report which design choices actually help once the artifact is removed --
including a negative result for uncertainty-aware fusion that the circular reference had
hidden.
