\section{Results}
\label{sec:results}

We evaluate on the independent, human-verified sparse-view reference described in
Section~\ref{sec:experiments}: a per-scene object multiset enumerated directly from the
RGB frames of all five TUM RGB-D~\cite{sturm2012tumrgbd} paper-subset sequences,
authored independently of the SAM, CLIP, VGGT~\cite{wang2025vggt}, and fusion stages, and
then human-verified ($n=5$ scenes). The primary metric is \emph{object-label F1}
(\texttt{object\_label\_f1}): multiset precision/recall/F1 of fused-node labels against the
reference multiset, reported as the mean over the five scenes at $3/5/8/10$ input views.
All numbers in this section are measured against this independent reference unless stated
otherwise; the contrast with the earlier prediction-seeded pseudo-reference is reported in
Section~\ref{subsec:decirc}.

\subsection{Variant comparison on the independent reference}
\label{subsec:variants}

Table~\ref{tab:variant_f1} compares seven variants. The four lifting/fusion baselines
(\textsc{2d-only}, \textsc{geometry-only}, \textsc{semantic-lifting}, and the
no-uncertainty \textsc{graph-fusion}), an uninformed \textsc{fixed-shrink} control that
applies a constant merge-gate shrink with no per-node signal, the uncertainty-aware
\textsc{proposed} variant (rank-normalized, weight $0.3$), and our duplicate-aware
\textsc{graph-fusion-dedup}. On real, detector-grounded labels, \textsc{graph-fusion} is
the strongest of the prior fusion variants at every view count, and object-label F1
\emph{falls} as views are added for every cross-view fusion variant (\textsc{semantic-lifting},
\textsc{fixed-shrink}, \textsc{graph-fusion}, and \textsc{proposed}); the non-fusion references
(\textsc{2d-only}, \textsc{geometry-only}) instead plateau. More views supply more detections
of the same physical object, which the cross-view fusion does not collapse, so precision (and
net F1) degrade against a fixed reference. \textsc{graph-fusion-dedup} is the only variant
whose accuracy \emph{rises} toward the dense end and is best at every view count.

\begin{table}[t]
\centering
\caption{Object-label F1 on the independent, human-verified reference (mean over five TUM
RGB-D sequences, full multiset). Best per column in bold. \textsc{fixed-shrink} is an
uninformed control; \textsc{proposed} is the rank-normalized uncertainty-aware variant
($w{=}0.3$). On real labels, accuracy falls with more views for every cross-view fusion
variant except duplicate-aware fusion, which is best throughout.}
\label{tab:variant_f1}
\small
\resizebox{\linewidth}{!}{%
\begin{tabular}{lrrrr}
\toprule
Variant & v3 & v5 & v8 & v10 \\
\midrule
\textsc{2d-only}                       & 0.4897 & 0.4940 & 0.4820 & 0.4647 \\
\textsc{geometry-only}                 & 0.4578 & 0.4843 & 0.4834 & 0.4833 \\
\textsc{semantic-lifting}              & 0.4168 & 0.3326 & 0.2548 & 0.2148 \\
\textsc{fixed-shrink} (control)        & 0.5165 & 0.4849 & 0.4471 & 0.4008 \\
\textsc{graph-fusion} (baseline)       & 0.5126 & 0.5038 & 0.4688 & 0.4324 \\
\textsc{proposed} (uncertainty, $w{=}0.3$) & 0.4982 & 0.4725 & 0.4132 & 0.3749 \\
\textbf{\textsc{graph-fusion-dedup} (ours)} & \textbf{0.5804} & \textbf{0.5923} & \textbf{0.6137} & \textbf{0.5787} \\
\bottomrule
\end{tabular}
}
\end{table}

\subsection{Duplicate-aware fusion: the positive result}
\label{subsec:dedup}

Duplicate-aware fusion adds a single post-fusion step: same-label fused nodes whose
axis-aligned 3D boxes overlap (IoU $> \tau$, default $\tau{=}0.1$) are merged and re-fused.
Table~\ref{tab:dedup_delta} reports the gain over both the \textsc{graph-fusion} baseline
and the uninformed \textsc{fixed-shrink} control, together with per-scene win/loss counts.
Dedup improves object-label F1 at every view count by $+0.068$ to $+0.146$ over
\textsc{graph-fusion} and by $+0.064$ to $+0.178$ over \textsc{fixed-shrink}, and the gain
\emph{grows} with views, as expected for a step that removes the duplicate detections that
accumulate as more frames are seen. It wins on every scene with no losses at the dense view
counts ($5\text{W}/0\text{L}$ at v8 and v10); the conservative sign-test floor at $n=5$ is
$p=0.062$.

\begin{table}[t]
\centering
\caption{Duplicate-aware fusion versus the \textsc{graph-fusion} baseline and the
uninformed \textsc{fixed-shrink} control. $\Delta$ is the object-label F1 improvement; W/L
counts are per-scene wins/losses for dedup over \textsc{graph-fusion} ($n=5$). The gain
grows with views and there are no losses at v8/v10.}
\label{tab:dedup_delta}
\small
\resizebox{\linewidth}{!}{%
\begin{tabular}{lrrrr}
\toprule
 & v3 & v5 & v8 & v10 \\
\midrule
$\Delta$ dedup $-$ \textsc{graph-fusion}  & $+0.068$ & $+0.089$ & $+0.145$ & $+0.146$ \\
$\Delta$ dedup $-$ \textsc{fixed-shrink}  & $+0.064$ & $+0.107$ & $+0.167$ & $+0.178$ \\
Per-scene W/L (vs \textsc{graph-fusion})  & 4W/0L & 4W/0L & 5W/0L & 5W/0L \\
\bottomrule
\end{tabular}
}
\end{table}

\subsection{Uncertainty-aware fusion: a negative result}
\label{subsec:uncertainty}

We treat the uncertainty-aware fusion gate --- the headline of an earlier framing of this
work --- as an ablation. Table~\ref{tab:uncertainty_neg} reports its delta against
\textsc{graph-fusion}. The recalibrated \textsc{proposed} variant is \emph{worse} at every
view count ($-0.014$ to $-0.057$) and loses on every scene at v5, v8, and v10
($0\text{W}/5\text{L}$). A sweep over the uncertainty weight $w \in \{0.1, 0.2, 0.3, 0.5,
0.8\}$ does not rescue it: \emph{no} weight in $[0.1, 0.8]$ beats \textsc{graph-fusion} at
any view count, and the best \textsc{proposed} setting still trails the baseline by
$-0.018/-0.052/-0.078/-0.082$ (v3/v5/v8/v10), with larger weights worse at the dense end.
The result is robust to reference filtering: dropping \texttt{unknown object} (which the
open-vocabulary detector never predicts) or the ``stuff'' classes (wall/floor/ceiling)
raises every variant's F1 but leaves the ranking unchanged.

\begin{table}[t]
\centering
\caption{Uncertainty-aware fusion is a negative result. $\Delta$ \textsc{proposed} $-$
\textsc{graph-fusion} in object-label F1 (rank-normalized, $w{=}0.3$); W/L is per-scene over
$n=5$. The variant is worse at every view count, and a weight sweep over $[0.1, 0.8]$
contains no setting that beats the baseline.}
\label{tab:uncertainty_neg}
\small
\resizebox{\linewidth}{!}{%
\begin{tabular}{lrrrr}
\toprule
 & v3 & v5 & v8 & v10 \\
\midrule
$\Delta$ \textsc{proposed} $-$ \textsc{graph-fusion} & $-0.014$ & $-0.031$ & $-0.056$ & $-0.057$ \\
Per-scene W/L                                         & --       & 0W/5L    & 0W/5L    & 0W/5L \\
\bottomrule
\end{tabular}
}
\end{table}

The mechanism is informative rather than a mere null. The uncertainty modulation does
exactly what it was designed to do --- it lifts recall by declining to merge uncertain
proposals (full-multiset recall, mean over scenes: \textsc{proposed} $0.505/0.567/0.645/0.643$
vs.\ \textsc{graph-fusion} $0.495/0.517/0.605/0.594$) --- but the same behavior fragments
correct objects and costs more precision than the recall it buys (precision: \textsc{proposed}
$0.512/0.421/0.314/0.272$ vs.\ \textsc{graph-fusion} $0.567/0.544/0.413/0.374$). Against a fixed-count reference
the net F1 is lower. Crucially, this is the \emph{opposite} axis to duplicate-aware fusion:
the uncertainty gate \emph{tightens} merging and splits more, whereas the system's measured
weakness (Section~\ref{subsec:perclass}) is over-counting, which calls for the opposite move
--- merging more. The same per-class evidence thus explains why uncertainty fails and why
dedup succeeds.

\subsection{Per-class characterization: precision-up, recall-down}
\label{subsec:perclass}

Table~\ref{tab:perclass_overall} reports the micro-averaged precision/recall/F1 of
\textsc{graph-fusion} per view count (dropping \texttt{unknown object}, which the detector
never emits). The signature is clear: recall rises from $0.671$ at three views to $0.808$
at ten, while precision falls from $0.533$ to $0.322$, so net F1 is best at the sparse end
($0.594$ at v3, $0.461$ at v10). The system finds more true objects with more frames but
emits more duplicate detections, and the precision loss dominates.

\begin{table}[t]
\centering
\caption{Per-class characterization of \textsc{graph-fusion} (micro-averaged over five
scenes, dropping \texttt{unknown object}). Recall rises and precision falls as views are
added; F1 is best at the sparse end. The precision loss is over-counting of
correctly-found objects, not mislabeling.}
\label{tab:perclass_overall}
\small
\resizebox{\linewidth}{!}{%
\begin{tabular}{rrrr}
\toprule
Views & Precision & Recall & F1 \\
\midrule
3  & 0.533 & 0.671 & \textbf{0.594} \\
5  & 0.464 & 0.699 & 0.557 \\
8  & 0.368 & 0.822 & 0.508 \\
10 & 0.322 & 0.808 & 0.461 \\
\bottomrule
\end{tabular}
}
\end{table}

Per class, this splits into three behaviors. A group is \emph{found reliably but
over-counted} --- recall saturates at $1.0$ while precision collapses with views: \texttt{desk}
($0.25 \to 0.17$), \texttt{keyboard} ($0.56 \to 0.26$), \texttt{monitor} ($0.55 \to 0.33$),
and \texttt{wall} ($1.00 \to 0.36$). These are always detected, but the 3D fusion does not
collapse the several detections of one physical object, making them the dominant precision
sink. A second group is \emph{clean wins} with high precision and recall (\texttt{book},
\texttt{bag}, \texttt{chair}, \texttt{plant}, and \texttt{cup}/\texttt{bottle} at sparse
views): distinct, well-separated objects that fuse correctly. A third group is \emph{genuine
misses}: \texttt{cabinet} (never detected) and \texttt{door} (recall $\approx 0.33$), where
the detector under-fires on blurry, low-resolution TUM frames. The dominant error is thus
over-counting of correctly-found objects, not mislabeling --- which is precisely what
duplicate-aware fusion targets.

Table~\ref{tab:dedup_mechanism} confirms the mechanism at v8: dedup sharply raises precision
on the over-counted classes (\texttt{monitor} $0.38 \to 1.00$, \texttt{desk} $0.19 \to 0.83$,
\texttt{keyboard} $0.31 \to 0.62$) while leaving recall unchanged for almost every class.
The one honest cost is closely-packed same-class instances: a desk holding two or three
monitors yields overlapping axis-aligned 3D boxes that box-IoU merges into one, costing
\texttt{monitor} $-0.33$ and \texttt{desk} $-0.17$ recall. The net remains a large F1 gain
because the precision recovery dwarfs the recall cost; separating closely-packed same-class
instances by appearance is left to future work (Section~\ref{sec:discussion}).

\begin{table}[t]
\centering
\caption{Per-class mechanism of duplicate-aware fusion at v8 (micro-averaged over scenes).
Dedup recovers precision on the over-counted classes while preserving recall for almost all
classes; the only recall cost is on closely-packed same-class instances (\texttt{monitor},
\texttt{desk}).}
\label{tab:dedup_mechanism}
\small
\resizebox{\linewidth}{!}{%
\begin{tabular}{lcccc}
\toprule
Class & \textsc{graph-fusion} P/R & dedup P/R & $\Delta$ recall \\
\midrule
\texttt{monitor}  & 0.38 / 1.00 & 1.00 / 0.67 & $-0.33$ \\
\texttt{desk}     & 0.19 / 1.00 & 0.83 / 0.83 & $-0.17$ \\
\texttt{keyboard} & 0.31 / 1.00 & 0.62 / 1.00 & $0.00$ \\
\texttt{wall}     & 0.40 / 1.00 & 1.00 / 1.00 & $0.00$ \\
\texttt{floor}    & 0.57 / 0.80 & 1.00 / 0.80 & $0.00$ \\
\texttt{book}     & 0.64 / 0.82 & 1.00 / 0.82 & $0.00$ \\
\texttt{chair}    & 0.50 / 0.80 & 0.80 / 0.80 & $0.00$ \\
\texttt{plant}    & 0.33 / 1.00 & 1.00 / 1.00 & $0.00$ \\
\bottomrule
\end{tabular}
}
\end{table}

\subsection{Robustness of the duplicate-aware merge}
\label{subsec:dedup_robust}

The dedup gain is robust to the IoU threshold. Table~\ref{tab:dedup_iou} sweeps
$\tau \in \{0, 0.05, 0.1, 0.15, 0.2, 0.3, 0.5\}$: the improvement over \textsc{graph-fusion}
stays in the $+0.086$ to $+0.123$ range (average $\Delta$) across the entire low-IoU band
$\tau \in [0, 0.2]$ and only fades once the threshold is strict enough to stop merging
duplicates ($\tau \ge 0.5$, $+0.026$). The wired default $\tau{=}0.1$ is within noise of the
optimum and is slightly more recall-conservative than ``any overlap.'' Adding a
centroid-distance OR-term on top of IoU does not help (v8: $0.653 \to 0.647/0.651$ for
distance $< 0.1/0.2$), so 3D-box IoU alone is the right and simplest criterion.

\begin{table}[t]
\centering
\caption{IoU-threshold sweep for duplicate-aware fusion (object-label F1, mean over five
scenes). The gain over \textsc{graph-fusion} is robust across the low-IoU band
$[0, 0.2]$ and only fades for strict thresholds. Default $\tau{=}0.1$ in bold.}
\label{tab:dedup_iou}
\small
\resizebox{\linewidth}{!}{%
\begin{tabular}{lrrrrr}
\toprule
IoU $\tau$ & v3 & v5 & v8 & v10 & avg $\Delta$ \\
\midrule
0.0 (any overlap) & 0.588 & 0.587 & 0.631 & 0.603 & $+0.123$ \\
0.05              & 0.580 & 0.585 & 0.607 & 0.585 & $+0.110$ \\
\textbf{0.1 (default)} & \textbf{0.580} & \textbf{0.592} & \textbf{0.614} & \textbf{0.579} & \textbf{$+0.112$} \\
0.15              & 0.577 & 0.589 & 0.597 & 0.558 & $+0.100$ \\
0.2               & 0.566 & 0.576 & 0.581 & 0.548 & $+0.088$ \\
0.3               & 0.552 & 0.551 & 0.537 & 0.516 & $+0.060$ \\
0.5               & 0.524 & 0.523 & 0.502 & 0.471 & $+0.026$ \\
\bottomrule
\end{tabular}
}
\end{table}

\subsection{De-circularization: the earlier pseudo-reference was misleading}
\label{subsec:decirc}

The two findings above only become visible under an independent reference. The earlier
evaluation of this system used a prediction-seeded pseudo-reference, drafted from the model's
own ten-view \textsc{graph-fusion} output. Because that reference encoded \textsc{graph-fusion}'s
own splitting pattern, two artifacts followed. First, \textsc{graph-fusion} scored $1.0$ at
ten views \emph{by construction} --- the reference \emph{was} its own prediction, so the
dense-view row measured reference construction rather than accuracy. Second, the
uncertainty-aware \textsc{proposed} variant, which keeps more uncertain proposals unmerged
and therefore splits more, reproduced more of the same fragments the circular reference
contained, and so appeared to be a sparse-view \emph{win} ($+0.040$ at v3, $+0.021$ at v5).
The independent reference fixes the true object count: over-splitting is now correctly
penalized, the uncertainty result reverses sign (Table~\ref{tab:uncertainty_neg}), and the
dense-view $1.0$ artifact disappears (F1 instead falls with views,
Table~\ref{tab:perclass_overall}). We document this pitfall because a prediction-seeded
reference is an easy and silent way to inflate scores in 3D scene understanding, and because
it is what masked both of this paper's main empirical findings.

\subsection{Larger-scale corroboration}
\label{subsec:expansion}

\todo{$\sim$30-scene corroboration pending. The headline variant comparison
(Table~\ref{tab:variant_f1}), the dedup gain (Table~\ref{tab:dedup_delta}), and the
uncertainty negative result (Table~\ref{tab:uncertainty_neg}) are established on $n=5$
human-verified scenes (consistent 5/5 at v5--v10, sign-test floor $p=0.062$). A
$\sim$30-scene expansion is in progress to raise statistical power; results slot directly
into the tables above. All quantitative claims in this paper should be read as $n=5$ until
this subsection is populated.}
