\documentclass[11pt]{article}

\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\usepackage{parskip}

\begin{document}

\begin{flushright}
Tanvir Ahmed\\
North South University\\
Dhaka, Bangladesh\\
\href{mailto:tanvir.ahmed32@northsouth.edu}{tanvir.ahmed32@northsouth.edu}\\
June 6, 2026
\end{flushright}

Editors-in-Chief\\
\emph{Computer Vision and Image Understanding}

\textbf{Re: Submission of ``Uncertainty-Aware Language-Grounded 3D Scene Graphs from Sparse RGB Views''}

Dear Editors-in-Chief,

Please consider our manuscript entitled ``Uncertainty-Aware Language-Grounded 3D Scene Graphs from Sparse RGB Views'' for publication as a Research Article in \emph{Computer Vision and Image Understanding}.

The manuscript addresses sparse-view image understanding: given only a small number of RGB views, the proposed system constructs an object-level, language-grounded 3D scene graph. The method combines feed-forward VGGT geometry, SAM mask proposals, CLIP and DINOv2 descriptors, uncertainty-aware 2D-to-3D lifting, cross-view object fusion, open-vocabulary node labeling, and spatial relation inference.

The paper contributes a sparse-view 3D scene graph pipeline, a reproducible five-sequence TUM RGB-D benchmark protocol, and a checked sparse-view consistency evaluation. The benchmark includes generated tables, figures, review packets, and checked prediction-seeded object/relation references. The evaluation shows monotonic improvements from 3 to 8 views in both object-label F1 and relation-triplet F1.

The paper fits CVIU's scope because it studies high-level interpretation of pictorial information and produces a structured representation of objects and relations from sparse RGB evidence. The evaluation is framed as checked consistency against reviewed 10-view references, and the manuscript explicitly distinguishes these metrics from independent dense-scene ground-truth accuracy.

The manuscript is original, has not been published previously, and is not under consideration elsewhere. It has been prepared for double-anonymized review. A separate title page, highlights file, data availability statement, declaration of competing interest, and generative-AI disclosure are included with the submission materials.

Sincerely,

\vspace{1em}
Tanvir Ahmed

\end{document}
